Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-11-28 submit/fix.tex
forging on

@@

In this article we use the idea that spectral indicators carry information on intrinsic supernova colors at peak magnitude.
This approach is taken by \citet{2011A&A...529L...4C}, who find that after standardization based on Ca and Si features, remaining residual color
variation is consistent with Milky Way dust models.
We accommodate up to  three independent color parameters: the model is constructed such that 
two are assigned as extrinsic processes attributed to dust, and one is attributed to intrinsic supernova
properties.
The data used in this analysis are described in \S\ref{data:sec}.
\replaced{
In \S\ref{model:sec} we present a
first analysis with two extrinsic parameters and omitting the intrinsic parameter.  The properties of the two extrinsic parameters
are consistent with the expectations of parameterized Milky Way extinction models.
In \S\ref{model2:sec} we add the intrinsic parameter in the analysis, and find strong evidence for its influence on observed magnitudes.
}{wefwe}
In  \S\ref{discussion:sec}
we relate our results to known
signatures of SN~Ia diversity.
Avoid repetitiiveness.
Conclusions are presented in \S\ref{conclusions:sec}.

@@

spectroscopic-feature measurements, SALT light-curve fit parameters including the shape  $x_1$,
and uncertainties

@@

\section{Model~I: Extrinsic Colors}
\label{modelI:sec}
We begin by considering a model for supernova magnitudes that have linear dependence on spectral parameters
and two extrinsic color parameters. This model is referred to as Model~I.  It is not the primary focus of this article, but its
results are of interest.  The purpose of this Section is to introduce some basic concepts of our models,
to show how the models can be expressed in terms of dust-extinction models, and to give Model~I results on dust. 
We refrain from presenting the full results of Model~I in deference to the more complex Model~II that is the subject of Section~\ref{modelII:sec},

\subsection{Model}
We assume 
that  peak intrinsic ${\hat{U}}{\hat{B}}\hat{V}\hat{R}\hat{I}$ magnitudes of a supernova are linearly dependent
on its
 equivalent widths of the Ca~II H\&K and Si~II~$\lambda$4141 spectral features
$EW_{Ca}$ and $EW_{Si}$,
and the wavelength of the minimum of 
the Si~II~$\lambda6355$ feature $\lambda_{Si}$
around $B$-band peak brightness:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
The explicit omission of light-curve shape in our model is compensated by its proxy,
$EW_{Si}$, at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
A grey magnitude offset, $\Delta$, is included for each supernova
to capture 
potential
band-independent intrinsic dispersion, while also absorbing peculiar-velocity errors introduced when converting
fluxes to luminosities.
The intrinsic  magnitudes are
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta \vec{I} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} \\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si}  \\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} \\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} \\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}
\end{pmatrix},
\label{ewsiv:eqn}
\end{equation}
where $\vec{I}$ is the Identity vector.
The global parameter vectors\footnote{Global parameters act on each of the five bands: a single component is written with a subscript for the corresponding band, e.g.\  $c_{\hat{U}}$; 
the combination of all components are written as a vector, e.g.\ $\vec{c}=(c_{\hat{U}}, c_{\hat{B}}, c_{\hat{V}}, c_{\hat{R}}, c_{\hat{I}}) )$,.  This 
notationally distinguishes between global versus per-supernova parameters.
} that describe the SN~Ia population
$\vec{c}$, 
$\vec{\alpha}$, $\vec{\beta}$,  and $\vec{\eta}$,  are the intercept
(units of mag) and slopes (units of mag~\AA$^{-1}$)
of the linear relationships that
relate spectral-feature parameters with intrinsic magnitudes.

@@

Except where stated otherwise above, we adopt a flat prior for all parameters.

@@

\subsection{Results: $\gamma^0$ and $\gamma^1$}
\label{results1:sec}
In this subsection we concentrate on the results for $\vec{\gamma}^0$ and $\vec{\gamma}^1$
(the 68\% credible intervals for all global parameters are given in \S\ref{m12results:sec}.)
In our model, $g_0$ and $g_1$ are introduced as latent parameters that are related to linear shifts in band magnitudes through
the  coefficients $\vec{\gamma}^0$ and
$\vec{\gamma}^1$.  Otherwise, no physically-motivated suppositions on the properties of $g$ and $\vec{\gamma}$ are imposed.
Attenuation of supernova light due to dust is expected to be mostly captured by the first-order affects described by our linear model.
Here we  interpret our model $\gamma$'s in terms of
the dust-extinction model of \citetalias{1999PASP..111...63F}.

All elements of the two $\vec{\gamma}$ parameter vectors 
describing intrinsic color
are significantly non-zero.
None of the 20000 links of 
our Monte Carlo chains for $\vec{\gamma}$ come close to the origin, we thus claim detection of the
influence of $g_0$ and $g_1$  on supernova magnitudes
with probability $1-5\times 10^{-5}$.
We thus find a strong signal for a linear 
extrinsic contribution to our model
\begin{equation}
A_X = \gamma^0_X g_0 +  \gamma^1_X g_1.
\end{equation}

A linear model with parameters $A_V$ and $E(B-V)$
\begin{equation}
A_X = a_X  A_{V} + b_X E(B-V)
\label{f99:eqn}
\end{equation}
approximately describes  
dust-extinction
models.  For example,   the wavelength-dependent model of  \citet{1989ApJ...345..245C} is linear,
though its linearity is lost when integrated over broad-band filters.
For the case of
$R^F=2.5$ and $A^F_V=0.1$ dust attenuating light from the SALT2
\citep{2007A&A...466...11G} $s=1$, $x_1=0$ SN~Ia template at ${\hat{B}}$-band peak, the 
\citet{1999PASP..111...63F} \citepalias[henceforth referred to as][]{1999PASP..111...63F} model
can be approximated with
$\vec{a} = (0.96,   1.00,   1.00,   0.97,   0.77)$ and $\vec{b}=(  1.77,   0.98,   0.12,  -0.50,  -0.53)$.
(The $F$ superscript is used to distinguish parameters of the \citetalias{1999PASP..111...63F} model.)
Over the ranges
 $0\le A^F_V\le 1$ and $2 \le R^F \le 3.5$ 
and the wavelengths under consideration here,  the values of the elements of $a$ and $b$ vary by $<5$\%
for the
\citetalias{1999PASP..111...63F} model with
 the largest error in extinction from the linear approximation having amplitude $<0.008$ mag.

Our model $\gamma$ vectors can be written as a linear combination (described by the $2 \times 2$ matrix M) of the
dust-based $a_X$ and $b_X$ vectors plus residual vectors perpendicular to $\vec{a}$ and $\vec{b}$,
\begin{equation}
\begin{pmatrix}
\vec{\gamma}^0 \\
\vec{\gamma}^1
\end{pmatrix}=
M
\begin{pmatrix}
\vec{a} \\
\vec{b}
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_{\gamma^0} \\
\vec{\epsilon}_{\gamma^1}
\end{pmatrix}.
\label{trans_I:eqn}
\end{equation}
The $\gamma$'s are almost entirely captured by the first term. 
The Model~I posterior gives a quadratic fraction of $0.9993^{+0.0002}_{-0.0002}$ of $\vec{\gamma}^0$ projected onto the plane defined by
$a_X$ and $b_X$, the quadratic fraction of $\vec{\gamma}^1$ projected
onto the dust plane is
$0.959^{+0.013}_{-0.024}$.
This result indicates that
the allowed variations in ${\hat{U}}{\hat{B}}{\hat{V}}{\hat{R}}{\hat{I}}$ allowed by the \citetalias{1999PASP..111...63F} model and our best-fit model are confined to almost identical
2-dimensional
planes within
the 5-dimensional magnitude space.
With no prior assumptions of dust extinction behavior or the distribution of $A_V$, the supernova data themselves exhibit
2-dimensional color variations that are closely aligned with the 2-dimensional color variations predicted by the \citetalias{1999PASP..111...63F} dust model.

@@

We
now turn to
the specific choice of $\vec{\gamma}$  returned by the fit within the context of the \citetalias{1999PASP..111...63F}  dust-extinction model. 
Our linear model would be satisfied by any two independent vectors  that span
the plane defined by the $\vec{\gamma}$ vectors.  While
Eq.~\ref{f99:eqn} expresses that model using basis vectors $\vec{a}$ and $\vec{b}$ and
their corresponding parameters 
$A_V$ and 
$E(B-V)$, the linear dust model 
alternatively
could be written in terms of any two other basis vectors  $\vec{a}+\kappa_1 \vec{b}$
and $\kappa_2 \vec{a} + \vec{b}$ as parameterized by $\kappa_1$ and $\kappa_2$.
In this basis, the model is then
\begin{equation}
A_X =  (1-\kappa_1 \kappa_2)^{-1} [(A^F_V - \kappa_2 E^F(B-V))\left(a_X+\kappa_1 b_X \right) +  (-\kappa_1 A^F_V + E^F(B-V)) (\kappa_2 a_X + b_X)],
\label{newdust:eqn}
\end{equation}
with the new parameters $A^F_V - \kappa_2 E^F(B-V)$ and $-\kappa_1 A^F_V + E^F(B-V)$  equal to  linear combinations of the original parameters.
Comparing this equation with  Eq.~\ref{trans_I:eqn}, we see that $\kappa_1=M[1,2]/M[1,1]$, $\kappa_2=M[2,1]/M[2,2]$.

Our posterior
includes a prior in which the per-supernova parameters $g_0$ and  $g_1$ associated with $\gamma$ are uncorrelated.
The parameters $\kappa_1$ and $\kappa_2$ are set by this prior.
If the external parameters of our model were indeed described by the \citetalias{1999PASP..111...63F}  dust model, then the prior corresponds
to    $A^F_V - \kappa_2 E^F(B-V)$ and $-\kappa_1 A^F_V + E^F(B-V)$ being uncorrelated in the supernova sample.
A trivial scenario that produces non-correlation is one in which all supernovae have the same parameter value for $-\kappa_1 A^F_V + E^F(B-V)$.
All color changes would then be confined to the $a_X+\kappa_1 b_X$ direction. This scenario corresponds to the \citetalias{1999PASP..111...63F}  dust model restricted to a constant $R^F=\kappa_1^{-1}$.   We use this interpretation to assign  $\langle R^F_{\mathit{eff}}\rangle = \kappa_1^{-1}$ as
the
ensemble mean
effective total-to-selective extinction of our sample.
With our posterior we
define an effective $R^F_{\mathit{eff}} \equiv \kappa^{-1}_1=2.18^{+0.14}_{-0.16}$.
Figure~\ref{plane:fig} visualizes the near-perfect alignment of $\vec{\gamma}^0$ and  $\vec{a}+\vec{b}/2.18$.

In the above context, the direction of the $\vec{\gamma}^1$ vector is described by $\kappa_2$.
The prior that $g_0$ and  $g_1$ are uncorrelated leads  to  $\kappa_2$ being determined
by the sample 
distribution of $A^F_V$ and $E^F(B-V)$.  In this article, we do not further consider the direction of
$\vec{\gamma}^1$ but rather focus on the distribution of the per-supernova parameters.

@@

\subsection{Remarks}
Model~I is not of primary interest in this article.  However, there are  points to highlight from its results.

\begin{itemize}
\item Our model introduces two latent supernova parameters that can influence supernova colors with ``non-informative'' priors.  Fitting
of our model gives a space of possible color variations highly consistent with those expected from two-parameter dust-extinction models. 
This lends confidence to our modelling and analysis.
\item In Model~I, any  intrinsic color variation that remains after spectral corrections is allocated to two parameters.  Color variations beyond
those from two-parameter dust extinction contribute to those two parameters.   The effective $R^F$ we derive from our results
is susceptible to leakage from intrinsic colors not associated with dust.
\end{itemize}

\section{Model~II: Extrinsic--Intrinsic Colors}
\label{modelII:sec}
Model~II extends Model~I to include a third intrinsic color parameter.  This increase in the number of degrees of freedom accommodates
the optical depth of dust, one parameter to capture variation in dust properties, and
intrinsic color variability unrelated to dust.  In this section we  focus on describing the consequences of adding the new parameter.

\subsection{Model}
\label{modelIImodel:sec}
Model~II is the same as Model~I except for the 
addition of an intrinsic color parameter $p$.  The distribution of this parameter over all supernovae is
assumed to be drawn from a Normal distribtion
\begin{equation}
p \sim \mathcal{N}(0, \sigma_p).
\end{equation}
The influence of $p$ on supernova colors is described by the global parameters $\vec{\phi}$ (constrained to be a
unit vector)
so that the color offset of each supernova is  $\vec{\phi}p$.
Model~II reads as
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} + \phi_{\hat{U}} p\\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si}  + \phi_{\hat{B}} p\\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} + \phi_{\hat{V}}  p\\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} + \phi_{\hat{R}} p\\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}+ \phi_{\hat{I}}  p
\end{pmatrix},
\label{ewsiv2:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} g_0+\gamma^1_{{\hat{V}}} g_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} g_0 + \gamma^1_{{\hat{R}}} g_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} g_0+\gamma^1_{{\hat{I}}} g_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust2:eqn}
\end{equation}

The new term is placed in Eq.~\ref{ewsiv2:eqn} as an intrinsic supernova property.  It is mathematically equivalent to
put the term in the mean of Eq.~\ref{dust2:eqn} joining what we call extrinsic parameters.  Although we 
use ``intrinsic'' and ``extrinsic'' as descriptive labels for $\phi$ and $\gamma$ respectively, the model cannot identify
their physical origins.

The model contains the product of latent parameters
$\vec{\phi} p$, which leads to the degeneracy $\vec{\phi} \rightarrow -\vec{\phi}$, $p \rightarrow -p$.
No conditions are added to break this degeneracy as we want to explore the possibility that
$\sigma_p\phi$ is consistent with zero.  The degeneracy
in $\vec{\phi}$ is apparent later in the article, but it does not affect the relevant science that
comes with the combination $\vec{\phi} p$.

>>>>

\added{
Even two parameters are probably not enough to capture supernova color diversity.  SNe~Ia 
are affected by interstellar Milky Way-like dust  and by ``intrinsic'' color variations from the supernova itself and its surrounding circumstellar dust. Physical modeling
shows that the complexity of interstellar dust shouldn't be expected to be captured by a single parameter \citep{2015ApJ...807L..26G,
2017ApJ...836...13H}. Supernova models \citep[e.g.][]{2007ApJ...662..487W,2013MNRAS.436..333S,2014MNRAS.441..532D,2015MNRAS.454.2549B,
2017ApJ...846...58H} and circumstellar dust \citep{2005ApJ...635L..33W,
2008ApJ...686L.103G} produce color variations beyond those due to interstellar dust.  A third color
parameter is therefore expected.}

@@

In this article we use the idea that spectral indicators carry information on intrinsic supernova colors at peak magnitude.
This approach is taken by \citet{2011A&A...529L...4C}, who find that after standardization based on Ca and Si features, remaining residual color
variation is consistent with Milky Way dust models.
We accommodate up to  three independent color parameters: the model is constructed such that 
two are assigned as extrinsic processes attributed to dust, and one is attributed to intrinsic supernova
properties.
The data used in this analysis are described in \S\ref{data:sec}.
\added{
The analysis itself is presented as a series of three models of increasing complexity.
\S\ref{modelI:sec}  and  \S\ref{modelII:sec} present the first two models, providing a soft introduction to
our methodology and yielding results to be compared with those from the third model.
That ultimate model and its results are discussed in detail in \S\ref{modelIII:sec}. 
In \S\ref{modelI:sec} we present Model~I, 
a first analysis using spectral features to standardize SN colors
and two additional latent parameters to account for color diversity. 
For Model~II described in \S\ref{modelII:sec}, we add a third latent parameter to the analysis.
Light-curve shape is then added to constitute our final Model~III in \S\ref{modelII:sec}.}
Conclusions are presented in \S\ref{conclusions:sec}.

@@

spectroscopic-feature measurements, \added{SALT2 light-curve shape parameters  $x_1$,}
and their uncertainties

@@

\section{Model~I: Extrinsic Colors}
\label{modelI:sec}
\added{
We begin by considering Model~I, in which supernova magnitudes  have linear dependence on spectral parameters
and two extrinsic color parameters. The purpose of this section is to introduce some basic concepts of our models,
to show how some of the  model parameters can be expressed in terms of dust-extinction models, and to give Model~I results on dust extinction. 
To avoid repetitiveness,
we refrain from going into full detail on Model~I in deference to Model~III, which is presented in Section~\ref{modelIII:sec}
and is the focus of this article.}

\subsection{Model}
We assume 
that  peak intrinsic ${\hat{U}}{\hat{B}}\hat{V}\hat{R}\hat{I}$ magnitudes of a supernova are linearly dependent
on its
 equivalent widths of the Ca~II H\&K and Si~II~$\lambda$4141 spectral features
$EW_{Ca}$ and $EW_{Si}$,
and the wavelength of the minimum of 
the Si~II~$\lambda6355$ feature $\lambda_{Si}$
around $B$-band peak brightness:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
The explicit omission of light-curve shape in our model is compensated by its proxy,
$EW_{Si}$, at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
A grey magnitude offset, $\Delta$, is included for each supernova
to capture 
potential
band-independent intrinsic dispersion, while also absorbing peculiar-velocity errors introduced when converting
fluxes to luminosities.
\added{With this grey offset, the model standardizes colors without reference to an absolute magnitude.
The intrinsic  magnitudes are
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta \vec{I} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} \\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si}  \\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} \\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} \\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}
\end{pmatrix},
\label{ewsiv:eqn}
\end{equation}
where $\vec{I}$ is the Identity vector.
The global parameter vectors\footnote{Global parameters act on each of the five bands: a single component is written with a subscript for the corresponding band, e.g.\  $c_{\hat{U}}$; 
the combination of all components are written as a vector, e.g.\ $\vec{c}=(c_{\hat{U}}, c_{\hat{B}}, c_{\hat{V}}, c_{\hat{R}}, c_{\hat{I}}) )$,.  This 
notationally distinguishes between global versus per-supernova parameters.
}} that describe the SN~Ia population
$\vec{c}$, 
$\vec{\alpha}$, $\vec{\beta}$,  and $\vec{\eta}$,  are the intercept
(units of mag) and slopes (units of mag~\AA$^{-1}$)
of the linear relationships that
relate spectral-feature parameters with intrinsic magnitudes.

@@

\added{Except where noted above, we adopt a flat prior for all parameters.

@@

\subsection{Results -- $\vec{\gamma}^0$ and $\vec{\gamma}^1$:  Cofactors the of Extrinsic Color Parameters}
\label{results1:sec}
In this subsection we concentrate on the results for $\vec{\gamma}^0$ and $\vec{\gamma}^1$.
In our model, $g_0$ and $g_1$ are introduced as latent parameters that are related to linear shifts in band magnitudes through
the  coefficients $\vec{\gamma}^0$ and
$\vec{\gamma}^1$.  Otherwise, no physically-motivated suppositions on the properties of $g$ and $\vec{\gamma}$ are imposed.
Attenuation of supernova light due to dust is expected to be mostly captured by the first-order affects described by our linear model.
We  interpret our model $\gamma$'s in terms of
the dust-extinction model of \citet{1999PASP..111...63F} \citepalias[henceforth referred to as][]{1999PASP..111...63F}.

The 68\% credible intervals for $\gamma^0$ and $\gamma^1$ (and all other global parameters) are given in \S\ref{m12results:sec}.}
All elements of the two $\vec{\gamma}$ parameter vectors 
describing intrinsic color
are significantly non-zero.
None of the 20000 links of 
our Monte Carlo chains for $\vec{\gamma}$ come close to the origin, we thus claim detection of the
influence of $g_0$ and $g_1$  on supernova magnitudes
with probability $1-5\times 10^{-5}$.
We thus find a strong signal for a linear 
extrinsic contribution to our model
\begin{equation}
A_X = \gamma^0_X g_0 +  \gamma^1_X g_1.
\end{equation}

\added{Dust-extinction
models are also approximately described by a linear model, here 
using the commonly-used parameterization of $A_V$ and $E(B-V)$
\begin{equation}
A_X = a_X  A_{V} + b_X E(B-V).
\label{f99:eqn}
\end{equation}
}
For example,   the wavelength-dependent model of  \citet{1989ApJ...345..245C} is linear,
though its linearity is lost when integrated over broad-band filters.
For the case of
$R^F=2.5$ and $A^F_V=0.1$ dust attenuating light from the SALT2
\citep{2007A&A...466...11G} $s=1$, $x_1=0$ SN~Ia template at ${\hat{B}}$-band peak, the 
\citetalias{1999PASP..111...63F} model
can be approximated with
$\vec{a} = (0.96,   1.00,   1.00,   0.97,   0.77)$ and $\vec{b}=(  1.77,   0.98,   0.12,  -0.50,  -0.53)$.
(The $F$ superscript is used to distinguish parameters of the \citetalias{1999PASP..111...63F} model.)
Over the ranges
 $0\le A^F_V\le 1$ and $2 \le R^F \le 3.5$ 
and the wavelengths under consideration here,  the values of the elements of $a$ and $b$ vary by $<5$\%
for the
\citetalias{1999PASP..111...63F} model with
 the largest error in extinction from the linear approximation having amplitude $<0.008$ mag.

\added{Our model $\vec{\gamma}$ vectors can be written as a linear combination (described by the $2 \times 2$ matrix M) of the
dust-based $\vec{a}$ and $\vec{b}$ vectors plus residual vectors perpendicular to the plane spanned by $\vec{a}$ and $\vec{b}$,
\begin{equation}
\begin{pmatrix}
\vec{\gamma}^0 \\
\vec{\gamma}^1
\end{pmatrix}=
M
\begin{pmatrix}
\vec{a} \\
\vec{b}
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_{\gamma^0} \\
\vec{\epsilon}_{\gamma^1}
\end{pmatrix}.
\label{trans_I:eqn}
\end{equation}
The $\vec{\gamma}$'s are almost entirely captured by the first term. 
A quadratic fraction of $0.9993^{+0.0002}_{-0.0002}$ of $\vec{\gamma}^0$ is projected onto $\vec{a}$--$\vec{b}$ plane, as is
$0.959^{+0.013}_{-0.024}$.of $\vec{\gamma}^1$.}
The allowed color
variations in ${\hat{U}}{\hat{B}}{\hat{V}}{\hat{R}}{\hat{I}}$ allowed by the \citetalias{1999PASP..111...63F} model and our best-fit model are confined to almost identical
2-dimensional
planes within
the 5-dimensional magnitude space.
With no prior assumptions of dust extinction behavior or the distribution of $A_V$, the supernova data themselves exhibit
2-dimensional color variations that are closely aligned with the 2-dimensional color variations predicted by the \citetalias{1999PASP..111...63F} dust model.

@@

We
now turn to
the specific choice of $\vec{\gamma}$  returned by the fit within the context of the \citetalias{1999PASP..111...63F}  dust-extinction model. 
Our linear model would be satisfied by any two independent vectors  that span
the plane defined by the $\vec{\gamma}$ vectors.  While
Eq.~\ref{f99:eqn} expresses that model using basis vectors $\vec{a}$ and $\vec{b}$ and
their corresponding parameters 
$A_V$ and 
$E(B-V)$, the linear dust model 
alternatively
could be written in terms of any two other basis vectors  $\vec{a}+\kappa_1 \vec{b}$
and $\kappa_2 \vec{a} + \vec{b}$ as parameterized by $\kappa_1$ and $\kappa_2$.
In this basis, the model is then
\begin{equation}
A_X =  (1-\kappa_1 \kappa_2)^{-1} [(A^F_V - \kappa_2 E^F(B-V))\left(a_X+\kappa_1 b_X \right) +  (-\kappa_1 A^F_V + E^F(B-V)) (\kappa_2 a_X + b_X)],
\label{newdust:eqn}
\end{equation}
with the new parameters $A^F_V - \kappa_2 E^F(B-V)$ and $-\kappa_1 A^F_V + E^F(B-V)$  equal to  linear combinations of the original parameters.
\added{The color vectors in this equation are associated with those of  Eq.~\ref{trans_I:eqn} by assigning $\kappa_1=M[1,2]/M[1,1]$, $\kappa_2=M[2,1]/M[2,2]$.}

Our posterior
includes a prior in which the per-supernova parameters $g_0$ and  $g_1$ associated with $\gamma$ are uncorrelated.
The parameters $\kappa_1$ and $\kappa_2$ are set by this prior.
If the external parameters of our model were indeed described by the \citetalias{1999PASP..111...63F}  dust model, then the prior corresponds
to    $A^F_V - \kappa_2 E^F(B-V)$ and $-\kappa_1 A^F_V + E^F(B-V)$ being uncorrelated in the supernova sample.
A trivial scenario that produces non-correlation is one in which all supernovae have the same parameter value for $-\kappa_1 A^F_V + E^F(B-V)$.
All color changes would then be confined to the $a_X+\kappa_1 b_X$ direction. This scenario corresponds to the \citetalias{1999PASP..111...63F}  dust model restricted to a constant $R^F=\kappa_1^{-1}$.   We use this interpretation to assign  $\langle R^F_{\mathit{eff}}\rangle \equiv \kappa_1^{-1}$ as
the
ensemble mean
effective total-to-selective extinction of our sample.
\added{
The posterior yields $R^F_{\mathit{eff}} =2.18^{+0.14}_{-0.16}$.
Figure~\ref{plane:fig} visualizes the near-perfect alignment of $\vec{\gamma}^0$ and  $\vec{a}+\vec{b}/2.18$.

The direction of the $\vec{\gamma}^1$ vector is described by $\kappa_2$.
The prior that $g_0$ and  $g_1$ are uncorrelated leads  to  $\kappa_2$ being determined
by the
distribution of $A^F_V$ and $E^F(B-V)$ for our supernova sample.  In this article, we do not further consider the direction of
$\vec{\gamma}^1$ but will later focus on the distribution of the per-supernova $g_0$ and $g_1$ parameters.

@@

\subsection{Remarks}
There are two points to highlight from the analysis of
Model~I.
\begin{itemize}
\item This model introduces two latent supernova parameters that can influence supernova colors with ``non-informative'' priors.  Fitting
of our model gives a space of possible color variations highly consistent with those expected from two-parameter dust-extinction models. 
That we recover a physically-motivated  result lends confidence in the modeling and analysis.
\item In Model~I, any  intrinsic color variation that remains after spectral corrections is allocated to two parameters.  Color variations beyond
those from two-parameter dust extinction contribute to those two parameters.   The derived effective dust-parameter $R^F$ 
is susceptible to leakage from intrinsic colors unassociated with dust.
\end{itemize}

\section{Model~II: Extrinsic and Intrinsic Colors}
\label{modelII:sec}
Model~II extends Model~I to include a third intrinsic-color parameter.  This increase in the number of degrees of freedom accommodates
one parameter for the optical depth of dust, one parameter to capture variation in dust properties, and
another parameter for intrinsic color variability unrelated to dust.  In this section we  focus on describing the consequences of the addition of the
new parameter.

\subsection{Model}
\label{modelIImodel:sec}
Model~II is identical to Model~I except for the 
addition of an intrinsic color parameter $p$.  The distribution of this parameter over all supernovae is
assumed to be drawn from a Normal distribution with unit standard deviation.
The relative effect on each band is specified by the global parameters 
$\sigma_p\vec{\phi}$, where $\vec{\phi}$ is a unit vector.
Model~II reads as
\begin{equation}
p  \sim \mathcal{N}(0,1 ),
\end{equation}
\begin{equation}
\sigma_p  \sim \text{Cauchy}(0.1,0.1),
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta  \vec{I} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} + \sigma_p\phi_{\hat{U}} p\\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si}  + \sigma_p\phi_{\hat{B}} p\\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} + \sigma_p\phi_{\hat{V}}  p\\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} + \sigma_p\phi_{\hat{R}} p\\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}+ \sigma_p\phi_{\hat{I}}  p
\end{pmatrix},
\label{ewsiv2:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} g_0+\gamma^1_{{\hat{V}}} g_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} g_0 + \gamma^1_{{\hat{R}}} g_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} g_0+\gamma^1_{{\hat{I}}} g_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust2:eqn}
\end{equation}

The new term is placed in Eq.~\ref{ewsiv2:eqn} as an intrinsic supernova property.  It is mathematically equivalent to
put the term in the mean of Eq.~\ref{dust2:eqn} joining what we call extrinsic parameters.  Although we 
use ``intrinsic'' and ``extrinsic'' as descriptive labels for $\phi$ and $\gamma$ respectively, the model cannot distinguish between
their physical origins.

The model contains the product 
$\vec{\phi} p$, which has the degeneracy $\vec{\phi} \rightarrow -\vec{\phi}$, $p \rightarrow -p$.
No conditions are added to break this degeneracy as we want to explore the possibility that
$\sigma_p\vec{\phi}$ is consistent with zero.  The degeneracy
in $\vec{\phi}$ is apparent later in the article, but it does not affect the magnitude predictions.

For $N$ supernovae there are $8N$ observables.  There are $3N$ spectral parameters, each of
$\Delta$, $g_0$, $g_1$ contributes $N-1$ parameters, there are $N$ parameters for $p$,  and $5 \times 7$ global coefficients.
For $N=172$ supernovae, there are 1376 observables and 1236  parameters.
